Classless Java

### 2015-09-27 paper/sections/Formalization.tex
little note on updating multiple fields
Method typing
Wops, I was wrong last time,
static behave differetly in interfaces w.r.t. abstract classes.
added \Gamma(this) is method typing
Some work for formalization

@@

@@

\subsubsection{Auxiliary function: \textsf{mbody}}

$\textsf{mbody}(m,C)$ denotes the actual body of method $m$ that interface $C$ owns. It can either be defined originally in $C$ or in its supertypes, and then passed to $C$ via inheritance.

The body of a method $m$ contains the relevant information with respect to that method, like the type of $m$ as well as the modifier. Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.

\[ \inferrule{
  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
  m \notin M}
{ \textsf{mbody}(m,C) = \emptyset } \]

\[ \inferrule{
  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
  m \in M}
{ \textsf{mbody}(m,C) = m } \]

\[ \inferrule{
  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
  \overline{M} \} \\
  m \notin M
  }
{ \textsf{mbody}(m,C_0) = \textsf{fold}_\textsf{shadow}\cdot\textsf{map}_{\textsf{mbody,}m}\cdot\textsf{subset}(\overline{C}) } \]

\[ \inferrule[(ReviewedRule)]{
  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
  \methods \} \\
  \methods(\m)=\method
\\
\Aux{tops}(\overline{C})=\C_1\ldots\C_n  
  }
{ \mBody(\m,\C_0) = \textsf{replace}(\method, \shadow(\mBody(\m,C_1)\ldots\mBody(\m,C_n))
 } \]
\subsubsection{Auxiliary function: \Aux{tops}}

\[ \textsf{tops}(\Cs)=\{\ \C\in\Cs\ |\ \nexists \C'\in\Cs\setminus\C,\ \C' \subtype \C\ \} \]

\subsubsection{Auxiliary function: \textsf{shadow}}

The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.

%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence
\begin{equation*}
\begin{array}{ll}
\shadow()=\emptyset\\
\shadow(\methods)=\method &\mif\ \method \in \methods, \forall \method' \in \methods\setminus\method :  \method' \subtype  \method,\\
\mbox{ otherwise}\\
\shadow(\methods)=\conflicted\mh&\mif\ \shadow(\Aux{removeBodies}(\methods))=\mh\\
%\conflictError \shadow \_\\
\_\subtype\emptyset\\
%\mh\mbox{\Q@\{return \_;\}@}\shadow \mh & \mif\ \mh\shadow\mh'\\
\_\ \mh\mbox{\Q@\{return \_;\}@}\subtype \mh & \mif\ \mh\subtype\mh'\\

\T \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\

%\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\textsf{ or }body_2.\textsf{modifier}=\textbf{static}\\
%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in}  & \textsf{if }body_1=\emptyset\textsf{ or }body_1.\textsf{modifier}=\textbf{static}
\end{array}
\end{equation*}

@@

>>>>

\footnote{Future work: updating multiple fields in one method call,
  \texttt{with(T v)}}

@@

\subsubsection{Method Typing} 
\[ 
\inferrule 
{ }
{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
\quad \textsc{(T-Meth)}
\]

\[ 
\inferrule
{\overline{x}:\overline{T} \vdash e:S \\ S <: T_0}
{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
    I} \\\\ \Gamma \vdash \textbf{this}:I }
\quad \textsc{(T-MethBody)}
\]

\[ 
\inferrule
{IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\
\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
T_0 <: U_0 }
{T_0 \spc m(\overline{T} \spc \overline{x}); \text{ OK IN I} }
\quad \textsc{(T-MethExt)}
\]

\[ 
\inferrule
{\overline{x}:\overline{T} \vdash e:S \\ S <: T_0 \\ 
IT(I)=\text{interface } I \text{ extends } \overline{J} \text{\{...\}} \\
\forall i,\text{if \textsf{mtype}}(m,J_i) = \overline(T) \to U_0, \text{then }
T_0 <: U_0 }
{T_0 \spc m(\overline{T} \spc \overline{x}) \text{ \{ return } e;\} \text{ OK IN
    I} \\\\ \Gamma \vdash \textbf{this}:I }
\quad \textsc{(T-MethBodyExt)}
\]

@@

\subsubsection{Functional notation}
We allows functional notation for set of methods $\methods$, using the method name $\m$ as a key.
That is, we define $\methods(\m)=\method$ iff there is a unique $\method\in\methods$ whose name is $\m$.
For convenience, we define $\methods(\m)=\none$ otherwise.

\subsubsection{Auxiliary function:\mBody}

$\mBody(\m,\C)$ denotes the actual body of method $\m$ that interface $\C$ owns. It can either be defined originally in $\C$ or in its supertypes, and then passed to $\C$ via inheritance.

The body of a method $\m$ contains all the relevant information with respect to that method, like the type of $\m$ as well as the modifier.
We use internally a special modifier $\conflicted$ to denote the case of two methods with conflicting implementation.
%
%\[ \inferrule{
%  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
%  m \notin M}
%{ \textsf{mbody}(m,C) = \emptyset } \]
%
%\[ \inferrule{
%  IT(C) = \text{\emph{ann} interface } C \{ \overline{M} \} \\
%  m \in M}
%{ \textsf{mbody}(m,C) = m } \]
%
%\[ \inferrule{
%  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
%  \overline{M} \} \\
%  m \notin M
%  }
%{ \textsf{mbody}(m,C_0) = \textsf{fold}_\textsf{shadow}\cdot\textsf{map}_{\textsf{mbody,}m}\cdot\textsf{subset}(\overline{C}) } \]

\[ \inferrule[(ReviewedRule)]{
  IT(C_0) = \text{\emph{ann} interface } C_0 \text{ extends } \overline{C} \{
  \methods \} \\
  \methods(\m)=\method
%\\
%\Aux{tops}(\overline{C})=\C_1\ldots\C_n  
  }
{ \mBody(\m,\C_0) = \textsf{replace}(\method, %\shadow(\mBody(\m,C_1)\ldots\mBody(\m,C_n))
\shadow(m,\Aux{tops}(\Cs)) 
} \]
\subsubsection{Auxiliary function: $\tops(\Cs), \shadow(\m,\Cs),\override(\method,\method')$}

\tops{} leave only the ``needed'' superinterfaces, that is, the ones that are not transitivelly reachable by following another superinterface chain.

The \textsf{shadow} function takes two same methods (with the same name and types of arguments), and return the method which shadows the other during inheritance.
%exmples to motivate our design
%interface A{static String m(){return "A";}}
%interface C extends A{
%	default String dm(){
%		  this.m();//wrong in java
%		  A.m();
%		  C.m();//wrong in java
%		}
%}
%
%
%(1) Static methods are not inherited. Also, if one of $\{body_1,body_2\}$ is null, \textsf{shadow} simply returns the other one. Hence

\begin{equation*}
\begin{array}{ll}
\tops(\Cs)=\{\ \C\in\Cs\ |\ \nexists \C'\in\Cs\setminus\C,\ \C' \subtype \C\ \} 
\\
\shadow(\m,\C_1\ldots\C_n)=\shadow(\methods)
& \mBody(\m,\C_i)\in\methods
\\ & \mif\ \mBody(\m,\C)=\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}
\\ &\mor\ \mBody(\m,\C)=\mh\QM;
\\

\shadow()=\none\\
\shadow(\methods)=\method &\mif\ \method \in \methods, \forall \method' \in \methods\setminus\method :  \method' \subtype  \method,\\
\mbox{ otherwise}\\
\shadow(\methods)=\conflicted\mh&\mif\ \shadow(\Aux{removeBody}(\methods))=\mh\\
\Aux{removeBody}(\QM{default}\ \mh\mbox{\Q@\{return \_;\}@})
=\mh\QM;\\
\Aux{removeBody}(\mh\QM)
=\mh\QM;
\\
%\mh\mbox{\Q@\{return \_;\}@}\shadow \mh & \mif\ \mh\shadow\mh'\\
\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\subtype \mh & \mif\ \mh\subtype\mh'\\

\T\ \m\oR\T_1\x_1\ldots \T_n\x_n\cR \subtype \T' \m\oR\T_1\x_1'\ldots\T_n\x_n'\cR & \mif\ \T\subtype \T'\\
\override(\method,\none)=
\override(\none,\method)=\method\\
\override(\method,\method')
=
\override(\method,\mh\QM;) & \method'=\QM{default}\ \mh\mbox{\Q@\{return \_;\}@}\\
\\
\override(\method,\mh')
=\method &\mif\ \method\in\{\mh\QM;,\QM{default}\ \mh\mbox{\Q@\{return \_;\}@} \}\\
& \mh\subtype\mh'\\
%\textsf{shadow}(body_1, body_2)=\emptyset & \textsf{if }body_1.\textsf{modifier}=body_2.\textsf{modifier}=\textbf{static}\\
%\textsf{shadow}(body_1, body_2)=body_1 & \textsf{if }body_2=\emptyset\textsf{ or }body_2.\textsf{modifier}=\textbf{static}\\
%\textsf{shadow}(body_1, body_2)=body_2 \hspace{.1in}  & \textsf{if }body_1=\emptyset\textsf{ or }body_1.\textsf{modifier}=\textbf{static}
\end{array}
\end{equation*}

@@

\subsubsection{Derived notations}

 Below shows how the functions $mtype$ and $mmodifier$ are derived from $mbody$.
